Stack Safety as a Security Property

### 2020-07-10 main.tex
CP 4

A key challenge in adapting the concepts of language-based security to this
setting is that integrity and confidentiality are ``nested'' properties: not
only does a given caller need to be protected from its callee, but the callee
also needs to be protected from {\em its} callees, and so on. That is, the notion
of ``attacker'' changes at many points throughout the execution of the program.
We address
this issue by introducing a dynamically changing notion of {\em accessibility} for
the different components of the machine state and adjusting the integrity
and confidentiality levels of different stack locations and registers at
every transition (call or return) between functions.

@@

The downside of the Depth Isolation policy is that it is rather
expensive to implement. In the same work,
\citeauthor{DBLP:conf/sp/RoesslerD18} propose a different mechanism
that is more efficient by virtue of being lazy in its enforcement.
Instead of enforcing the integrity and confidentiality of the stack at the
exact moment
when a callee returns, this policy signals violations when the
caller actually accesses data that the callee has written into its stack
frame\ifaftersubmission\bcp{We could also hint at the lazy confidentiality
  property inthe same way...}\fi.
The lazy policy does not enforce our original integrity and confidentiality
properties, and to our knowledge there is no formal statement of the properties that
it ought to enforce. We propose ``observable'' versions of integrity and
confidentiality property, which capture the desired behavior of a policy that
detects violations between when they occur and when they make a visible impact.
Unfortunately the existing Lazy Tagging and Clearing policy of
\citeauthor{DBLP:conf/sp/RoesslerD18} does not quite enforce these properties,
but would if enhanced (at some performance cost).
%\bcp{I worry that we are being a little
%  dishonest here: the stated advantage of the lazy policy is that it is more
%efficient, but our variant is {\em not} efficient!}

\sna{There's been some inconsistency in ``stack safety'' vs.
     ``stack-safety,'' I'm fixing on the former.}

In summary, we offer the following contributions:
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?}
  A new {\em formal stack safety property} capturing the
  intuitive concepts of strict local state encapsulation and well-bracketed
  control flow~(\cref{sec:lse-and-wbcf}) in terms of familiar security concepts.
  % These properties   rule out common examples of stack memory abuse.
  This property can be enforced by
  the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, which is
  based on hardware tagging~(\cref{sec:conservative}).
\item
  A {\em lockstep} variant of the stack safety property that is also
  enforced by the Depth Isolation policy
  but is more efficient to test, plus a proof (in Coq) that this variant is
  stronger than the original~(\cref{sec:lockstep}).
\item
  A {\em lazy} variant of the stack safety property that can be enforced if the
  Lazy Tagging and Clearing policy of \citet{DBLP:conf/sp/RoesslerD18} is enhanced
  with unique tags on each activation, plus a proof (in Coq) that this
  variant is weaker than the original~(\cref{sec:lazy}).
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

@@

\paragraph*{Capability machines}
%
%Other policies are implemented in hardware or built on hardware mechanisms.
Capability machines such as CHERI \citep{Woodruff+14} extend conventional
architectures to support efficient and fine-grained control over memory
accesses, which can be used to implement security policies, including
the stack protection policies of
\citet{Skorstengaard+19,Skorstengaard+19b}\bcp{I thought these policies used
  linear capabilities?}\leo{Only the second one AFAIK. What would you
  change to characterize it more accurately?}\rb{StkPointers uses linear
  capabilities, its predecessor users local capabilities (and unlike
  StkPointers could be implemented in CHERI)}.
Capabilities package up pointers
(ordinary data) with base and bounds information (policy data),
and they require rewriting source code (e.g., to use capabilities in place
of ordinary pointers), so separating ordinary machine state from policy
state is delicate; we conjecture it can be done with some effort.

@@

A trace over a type $\TRACEELEMS$ is a nonempty, finite or infinite sequence of
$\TRACEELEMS$s. We will mostly be concerned with sequences of pairs of machine states
and observations, called machine traces and denoted $\machT \in \MOTRACE$,
and with sequences of triples of a machine state, a policy state, and an
observation, denoted $\MPT \in \MPOTRACE$ and referred to as MP-traces.

@@

The ``trace-of'' operator, written \(\mach \hookrightarrow \machT\),
coinductively relates a machine state with the trace of machine states
paired with observations produced by repeated application of step:

\judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machT\)}
            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machT\)}%
%
\noindent
Since the step function is total, if \(\mach \hookrightarrow \machT\),
then \(\machT\) must be infinite.

In our running example in \Cref{fig:running-trace-b} we can
see a depiction of such a trace: starting from an initial machine
state at the very top, each instruction constructs a new machine state
(the diff of which compared to the previous one appears in the
``Effect'' column) and the majority of observations are $\tau$ with
the exception of instructions 31 and 13 which are writes to $\rout$
and where the observations are $0$ and $60$ respectively.

@@

\paragraph*{Until}
The operation \(\PUT~(f,\Trace)\)
%\bcp{Wouldn't
%``While'' be more mnemonic?} \sna{Appealing, but it's more of a
%``while-not''? Trying out ``until''. In any case... moving away
%from doing joins means this needs to get some work to eliminate
%the trailing observation. Here's an attempt.}
takes a trace
\(\Trace : \mathit{trace} ~ \TRACEELEMS \times \OBSS\) and a predicate
on elements \(f \subset \TRACEELEMS\) and gives the prefix of
\(\Trace\) ending with the first element on which \(f\) holds.
Subtly, because the observation associated with a trace represents its
next step, when we take a prefix from a trace containing observations
we must replace the final observation with a \(\tau\).

  \begin{center}
  \begin{minipage}{.3\textwidth}
    \judgment{\(f\ \traceelem\)}
             {\(\PUT~(f,(\traceelem,\obs)\Trace) = (\traceelem,\tau)\)}
  \end{minipage}
%
  \begin{minipage}{.3\textwidth}
  \judgment{}
           {\(\PUT~(f,(\traceelem,\obs)) = (\traceelem,\tau)\)}
%
  \end{minipage}
  \begin{minipage}{.3\textwidth}
  \judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT~(f,\Trace) = \Trace'\)}
              {\(\PUT~(f,(\traceelem,\obs)\Trace) = (\traceelem,\obs)\Trace'\)}
  \end{minipage}
  \end{center}
%
If the resulting trace is a strict prefix of the parameter, then the
predicate must hold on its final element. If the predicate never holds,
\(\PUT\) is the identity function.

@@

We will frequently take a prefix of the trace from an initial state up
through the first state where some condition holds on the machine state.
This can easily be implemented using \(\PUT\), and we provide special notation
for convenience.
%
Let \(f\) be a predicate on machine states. Then we define \(\mach
\hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix of
running \(\mach\) up to \(f\)''):

\judgmenttwo{\(\mach \hookrightarrow \machT'\)}
            {\(\PUT~(f,\machT')
              = \machT\)}
            {\(\mach \hookrightarrow \machT | f\)}
\noindent
And similarly for traces with policies:

@@

We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
as far as an external observer is concerned, if the sequence of
non-silent observations of $\obsT_1$ is a prefix of those of
$\obsT_2$; that is, we operate up to deletion of \(\tau\) observations,
coinductively:

@@

We then define similarity of observation traces as traces prefixing each other:

\[\obsT_1 \simeq \obsT_2 \triangleq \obsT_1 \lesssim \obsT_2 \land \obsT_2 \lesssim \obsT_1\]
%
%\leo{I still don't know how to do this transition. Maybe contrast
%with the ``perhaps surprisingly'' later?} Note that
%similarity of observation traces could alternatively be defined coinductively,
%just like the prefix relation, but by dropping the asymmetric rules that allow
%for a finite trace ($\tau \lesssim \obsT$ or $w \lesssim w\obsT$).\apt{Is this immportant?}\leo{I agree. Probably not}

Note that an infinite silent trace is a
prefix of (and equivalent to) any other trace. While this might seem
surprising at first, it makes sense in a timing-insensitive context:
an external observer looking at two machine runs cannot (computably)
distinguish between a machine that steps forever and a machine that
steps for a long time before producing some output.

>>>>

A key challenge in adapting the concepts of language-based security to this
setting is that stack safety is a ``nested'' property: not only does a given
caller need to be protected from its callee, but the callee also needs to be
protected from {\em its} callees, and so on. That is, the notion of
``attacker'' changes at many points throughout the execution of the program.
We address
this issue by introducing a dynamically changing notion of {\em accessibility} for
the different components of the machine state and adjusting the integrity
and confidentiality levels of different stack locations and registers at
every transition (call or return) between functions.

@@

The downside of the Depth Isolation policy is that it is rather
expensive to implement. In the same work,
\citeauthor{DBLP:conf/sp/RoesslerD18} propose a different mechanism
that is more efficient by virtue of being lazy in its enforcement.
Instead of enforcing the integrity and confidentiality of the stack at the
exact moment
when a callee returns, this policy signals violations when the
caller actually uses data that the callee has written into  its stack
frame\ifaftersubmission\bcp{We could also hint at the lazy confidentiality
  property inthe same way...}\fi.
The lazy
policy does not enforce our original stack-safety property, but (a slightly
enhanced variant of the policy\bcp{I worry that we are being a little
  dishonest here: the stated advantage of the lazy policy is that it is more
efficient, but our variant is {\em not} efficient!}) of it does enforce a slightly weaker
lazy version of the property that still protects from unsafe callees but is
less intuitive to state.

In summary, we offer the following contributions:
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
  A new {\em formal stack-safety property} capturing the
  intuitive concepts of strict local state encapsulation and well-bracketed
  control flow~(\cref{sec:lse-and-wbcf}) in terms of familiar security concepts.
  % These properties   rule out common examples of stack memory abuse.
  This property can be enforced by
  the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, which is
  based on hardware tagging~(\cref{sec:conservative}).
\item
  A {\em lockstep} variant of the stack-safety property that is also
  enforced by the Depth Isolation policy
  but is more efficient to test, plus a proof (in Coq) that this variant is
  stronger than the original~(\cref{sec:lockstep}).
\item
  A {\em lazy} variant of the stack-safety property that can be enforced by
  (a variant of\bcp{same concern}) the
  more efficient Lazy Tagging and Clearing policy
  of \citet{DBLP:conf/sp/RoesslerD18}, plus a proof (in Coq) that this
  variant is weaker than the original~(\cref{sec:lazy}).
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

@@

\paragraph*{Capability machines}
%
%Other policies are implemented in hardware or built on hardware mechanisms.
Capability machines such as CHERI \citep{Woodruff+14} extend conventional
architectures to support efficient and fine-grained control over memory
accesses, which can be used to implement security policies, including
the stack protection policies of
\citet{Skorstengaard+19,Skorstengaard+19b}\bcp{I thought these policies used
  linear capabilities?}\leo{Only the second one AFAIK. What would you
  change to characterize it more accurately?}.  Capabilities package up pointers
(ordinary data) with base and bounds information (policy data),
and they require rewriting source code (e.g., to use capabilities in place
of ordinary pointers), so separating ordinary machine state from policy
state is delicate; we conjecture it can be done with some effort.

@@

A trace over a type $\TRACEELEMS$ is a nonempty, finite or infinite sequence of
$\TRACEELEMS$s; we write $\TRACE{\TRACEELEMS}$ for the set of traces over
$\TRACEELEMS$. We will mostly be concerned with sequences of pairs of
machine states
and observations, called machine traces and denoted $\machT \in \MOTRACE$,
and with sequences of triples of a machine state, a policy state, and an
observation, denoted $\MPT \in \MPOTRACE$ and referred to as MP-traces.

@@

The ``trace-of'' operator, written \(\mach \hookrightarrow \machT\),
coinductively relates a machine state with the trace of machine states
and observations produced by repeated application of step:

\judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machT\)}
            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machT\)}%
%
\noindent
(We use juxtaposition of an element and a sequence to represent ``cons.'')
Since the step function is total, if \(\mach \hookrightarrow \machT\),
then \(\machT\) must be infinite.

In our running example in \Cref{fig:running-trace-b} we can
see a depiction of such a trace: starting from an initial machine
state at the very top, each instruction leads to a new machine state
(the diff of which, compared to the previous one, appears in the
``{\textit{Effect}'' column), and the majority of observations are $\tau$, except
for instructions 31 and 13, which are writes to $\rout$
where the observations are $0$ and $60$, respectively.

@@

\paragraph*{Until}
The operation \(\PUT~(f,\Trace)\)
%\bcp{Wouldn't
%``While'' be more mnemonic?} \sna{Appealing, but it's more of a
%``while-not''? Trying out ``until''. In any case... moving away
%from doing joins means this needs to get some work to eliminate
%the trailing observation. Here's an attempt.}
takes a trace\bcp{I changed a : here to $\in$.  We should check if there are
any more to fix.}
\(\Trace \in \TRACE{\TRACEELEMS} \times \OBSS\) and a predicate
on elements \(f \subset \TRACEELEMS\) and gives the prefix of
\(\Trace\) ending with the first element on which \(f\) holds.
One subtlety is that, because the observation associated with some element
in a trace
represents what can be observed when this element is executed (i.e., the
observation on that step), when we take a prefix from a trace containing
observations
we must replace the final observation with a \(\tau\).

  \begin{center}
  \begin{minipage}{.3\textwidth}
    \judgment{\(f\ \traceelem\)}
             {\(\PUT~(f,(\traceelem,\obs)\Trace) = (\traceelem,\tau)\)}
  \end{minipage}
%
  \begin{minipage}{.3\textwidth}
  \judgment{}
           {\(\PUT~(f,(\traceelem,\obs)) = (\traceelem,\tau)\)}
%
  \end{minipage}
  \begin{minipage}{.3\textwidth}
  \judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT~(f,\Trace) = \Trace'\)}
              {\(\PUT~(f,(\traceelem,\obs)\Trace) = (\traceelem,\obs)\Trace'\)}
  \end{minipage}
  \end{center}
%
If the resulting trace is a strict prefix of the parameter, then the
predicate must hold on its final element. If the predicate never holds, then
\(\PUT\) is the identity function.

@@

We will frequently take a prefix of the trace from an initial state up
through the first state where some condition holds on the machine state.
This can easily be implemented using \(\PUT\), and we provide special notation
for convenience.
%
If \(f\) be a predicate on machine states, we define \(\mach
\hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix of
running \(\mach\) up to \(f\)'')

\judgmenttwo{\(\mach \hookrightarrow \machT'\)}
            {\(\PUT~(f,\machT')
              = \machT\)}
            {\(\mach \hookrightarrow \machT | f\)}
\noindent
and similarly for traces with policies:

@@

We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
as far as an external observer is concerned, written \(\obsT_1 \lesssim
\obsT_2\), if the sequence of
non-silent observations of $\obsT_1$ is a prefix of those of
$\obsT_2$; that is, we operate up to deletion of \(\tau\) observations,
coinductively:

@@

\smallskip
We then define similarity of observation traces as traces prefixing each other:
\[\obsT_1 \simeq \obsT_2 \triangleq \obsT_1 \lesssim \obsT_2 \land \obsT_2 \lesssim \obsT_1\]
%
%\leo{I still don't know how to do this transition. Maybe contrast
%with the ``perhaps surprisingly'' later?} Note that
%similarity of observation traces could alternatively be defined coinductively,
%just like the prefix relation, but by dropping the asymmetric rules that allow
%for a finite trace ($\tau \lesssim \obsT$ or $w \lesssim w\obsT$).\apt{Is this immportant?}\leo{I agree. Probably not}

Note that an infinite silent trace is a
prefix of (and equivalent\bcp{similar?} to) any other trace. While this might seem
surprising at first, it makes sense in a timing-insensitive context:
an external observer looking at two machine runs cannot (computably)
distinguish between a machine that steps forever and a machine that
steps for a long time before producing some output.
